Scikit-learn: machine learning without learning the machinery

### 2014-11-06 article.tex
Fix typo (Java -> Weka)

\author[1]{G. Varoquaux\thanks{gael.varoquaux@inria.fr}}
\author[2]{L. Buitinck}
\author[3]{G. Louppe}
\author[1]{O. Grisel}
\author[1]{F. Pedregosa}
\author[4]{A. Mueller}
\affil[1]{Parietal, INRIA, Bat 145 CEA Saclay, 91191 Gif-sur-Yvette,
France}
\affil[2]{ILPS, University of Amsterdam, Science Park 904, Amsterdam, The
Netherlands}
\affil[3]{Systems and Modeling Research Unit, University of Liège
Montefiore Institute (B28), Liège, Belgium}
\affil[4]{Amazon Development Center Germany, Berlin, Germany}

@@

\paragraph{Project vision.}
%
Scikit-learn was born from the observation that most standard
machine learning algorithms were out of reach of the users that could
most benefit from them: researchers --biologists, climate
scientists, experimental physicists-- or developers of web
services or domain-specific applications.
%
In the scientific world, implementations of these algorithms
mostly consisted of scattered pieces of code %to download
on researchers' web pages. Unifying efforts could be found in
statistic-specific environments and libraries such as the R
language \cite{Rmanual}, the Weka Java toolkit \cite{hall2009weka}, or the
Shogun C++ library \cite{sonnenburg2010}.
Of these, R is a custom \emph{language},
while Java is a library with a custom GUI, and Shogun is a fairly
technical C++ code base.
In scikit-learn, we chose Python to reach many users, technical or not, tying machine learning
into a general-purpose language good for simple
 interactive programming \cite{perez2007ipython}, sophisticated
scripting, as well as full-blown applications.

@@

\paragraph{Some history.}
%
Scikit-learn started circa 2007 as code from David Cournapeau and
Matthieu Brucher's PhD work, but dwindled down till 2010, in which year the
Parietal team from INRIA adopted the project, hiring a full time
engineer. Basic APIs were defined and an efficient binding of LibSVM
\cite{chang2011libsvm} gave the project a compelling advantage. In
December 2011, the first international sprint was organized with generous
funding from Google. Today, the project has grown vastly beyond
INRIA into a worldwide open source effort with more than 200 contributors.

@@

\begin{figure*}[b]
% To have the float on the bottom, we need this declared on the page
% before where it is going to be printed
\lstinputlisting[lastline=12]{nlp_example.py}
\caption{Simple text-processing code to download movie reviews from
Internet and learn positive versus negative ratings.\label{fig:code}}
\end{figure*}

@@

Our development workflow relies on intensive code review,
facilitated by the online version control environment GitHub (\url{http://github.com})
The code review process is a great asset to
raising the quality of the code, from all points of view: algorithmic,
numerical robustness, homogeneity of APIs, ease of use, documentation.
Another central aspect for quality in the project is the heavy use of
unit testing (up to 88\% of the lines of code are covered), run
continuously on all contributions thanks to the Travis platform
(\url{http://travis-ci.org}).

@@

Scikit-learn has become a reference machine learning toolkit. As such, it
should offer a wide variety of learning algorithms. However, each new
feature, each new line of code, comes with a significant maintenance cost.
Uncontrolled growth would lead the project to pick up weight until it
came to a crawl. The difficulty lies in defining the project scope and
prioritizing methods. We strive to identify and integrate all major
methods most useful for prediction tasks. We favor time-tested
algorithms and put great care in the choice of default parameters. We put
a lot of effort in making the documentation didactic and
pragmatic. Our
secret ambition is that it can provide an executable alternative to
textbooks, with all the figures generated by code examples.

Machine learning is a very active field of research, with promising
trends such as deep learning, harnessing GPUs or distributed computing
for extra computational power, or extending to new platforms, such as
Android or iOS-based embedded devises. However, as scikit-learn purpose
is to provided tried-and-true, mostly turn-key, predictive tools, we have
to walk the fine line between supporting powerful techniques and
utilizing our limited project cycles on proven technologies. For
instance, deep neural networks are an emerging tool with a huge
potential, however, they are very much a technology in flux and currently
are far from being usable with a significant expertise. Similarly, GPUs
give a lot of user-level problems, as there is a lack of binary interface
to address them uniformly and robustly over all hardwares. Finally, we
support new platforms, such as Android, as long as they come with a good
Python implementation with a standard build chain. Indeed, our basic
execution environment is the scientific Python stack, that is both very
powerful and high-level. The vision of the scikit-learn project is to
bring advances in machine learning to users that that may lack the full
expertise. As a consequence, we focus on seasoned approaches, rather than
implementing ongoing research. We believe that code is an intrinsic part
of the scholarship outcome of scientific research in computational
science \cite{buckheit1995wavelab}, and scikit-learn strives to be akin
to a reference textbook in this research software landscape.

>>>>

\author[1]{G. Varoquaux\thanks{gael.varoquaux@inria.fr}}
\author[2]{L. Buitinck}
\author[3]{G. Louppe}
\author[1]{O. Grisel}
\author[1]{F. Pedregosa}
\author[4]{A. Mueller}
\affil[1]{Parietal, INRIA, Bat 145 CEA Saclay, 91191 Gif-sur-Yvette,
France}
\affil[2]{ISLA, University of Amsterdam, Science Park 904, Amsterdam, The
Netherlands}
\affil[3]{Systems and Modeling Research Unit, University of Liège
Montefiore Institute (B28), Liège, Belgium}
\affil[4]{Amazon Development Center Germany, Berlin, Germany}

@@

\paragraph{Project vision.}
%
Scikit-learn was born from the observation that most standard
machine learning algorithms were out of reach of the users that could
most benefit from them: researchers --biologists, climate
scientists, experimental physicists-- or developers of web
services or domain-specific applications.
%
In the scientific world, implementations of these algorithms
mostly consisted of scattered pieces of code %to download
on researchers' web pages. Unifying efforts could be found in
statistic-specific environments and libraries such as the R
language \cite{Rmanual}, the Weka Java toolkit \cite{hall2009weka}, or the
Shogun C++ library \cite{sonnenburg2010}.
Of these, R is a custom \emph{language},
while Weka is a Java library with a custom GUI, and Shogun is a fairly
technical C++ code base.
In scikit-learn, we chose Python to reach many users, technical or not, tying machine learning
into a general-purpose language good for simple
 interactive programming \cite{perez2007ipython}, sophisticated
scripting, as well as full-blown applications.

@@

\paragraph{Some history.}
%
Scikit-learn started circa 2007 as code from David Cournapeau and
Matthieu Brucher's PhD work, but dwindled down until 2010, when the
Parietal team at INRIA adopted the project, hiring a full-time
engineer. Basic APIs were defined and an efficient binding of LibSVM
\cite{chang2011libsvm} gave the project a compelling advantage. In
December 2011, the first international sprint was organized with generous
funding from Google. Today, the project has grown vastly beyond
INRIA into a worldwide open source effort with more than 200 contributors.

@@

\begin{figure*}[b]
% To have the float on the bottom, we need this declared on the page
% before where it is going to be printed
\lstinputlisting[lastline=12]{nlp_example.py}
\caption{Simple text-processing code to download movie reviews
and learn positive versus negative ratings.\label{fig:code}}
\end{figure*}

@@

Our development workflow relies on intensive code review,
facilitated by the online version control environment GitHub (\url{http://github.com})
The code review process is a great asset to
raising the quality of the code, from all points of view: algorithmic,
numerical robustness, homogeneity of APIs, ease of use, documentation.
Another central tool for maintaining quality in the project is the use of
unit testing (some 88\% of the lines of code are covered).
All would-be code contributions are automatically tested using Travis
(\url{http://travis-ci.org}).

@@

Scikit-learn has become a reference machine learning toolkit. As such, it
should offer a wide variety of learning algorithms. However, each new
feature, each new line of code, comes with a maintenance cost.
Uncontrolled growth would lead the project to pick up weight until it
came to a crawl. The difficulty lies in defining the project scope and
prioritizing methods. We strive to identify and integrate all major
methods most useful for prediction tasks. We favor time-tested
algorithms and put great care in the choice of default parameters. We put
a lot of effort in making the documentation didactic and
pragmatic. Our
secret ambition is that it can provide an executable alternative to
textbooks, with all the figures are generated by code examples.

Machine learning is a very active field of research, with promising
trends such as deep learning, harnessing GPUs or distributed computing
for extra computational power, or extending to new platforms, such as
Android or iOS-based embedded devises. However, as scikit-learn's purpose
is to provided tried-and-true, mostly turn-key, predictive tools, we have
to walk the fine line between supporting powerful techniques and
utilizing our limited project cycles on proven technologies. For
instance, deep neural networks are an emerging tool with a huge
potential, but are still in flux and hard to use without significant expertise.

Similarly, GPGPU technology is promising, but still poses user-level problems.
While OpenCL largely hides the incompatibility of GPUs and other compute devices
at the source code level \cite{stone2010opencl}, it does not guarantee performance portability
and requires the installation of device-specific drivers.
PyOpenCL \cite{klockner2012pycuda} may alleviate the former problem and communicates
with NumPy, making it an obvious starting point for experimentation
in the direction of GPU-enabled scikit-learn. Theano \cite{bergstra2010theano} has
shown the feasibility of programming GPU-powered neural nets in Python.

Support for mobile platforms such as Android is determined by those platforms'
support for Python and the standard C/C++ build chain. Indeed, our basic
execution environment is the scientific Python stack, that is both very
powerful and high-level. The vision of the scikit-learn project is to
bring advances in machine learning to users that that may lack the full
expertise. As a consequence, we focus on seasoned approaches, rather than
implementing ongoing research. We believe that code is an intrinsic part
of the scholarship outcome of scientific research in computational
science \cite{buckheit1995wavelab}, and scikit-learn strives to be akin
to a reference textbook in this research software landscape.
